Revisiting the Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation using complex Empirical Orthogonal Functions

### 2021-08-12 analysis/paper/shceof.Rmd
Arregla leyendas

theme_set(theme_shceof(base_size = 10))

@@

  geom_qmap(size = 0.1) +
  scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(10))  +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL, 
                   limits = c(NA, -20)) +
  scale_x_longitude() +
  geom_coords() +
  coord_polar() +
  theme(axis.text = element_blank()) +
  facet_wrap(.~EOF, ncol = 2)  +
  tagger::tag_facets("rc")

@@

```{r enso-phase, fig.cap = "ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit()

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) -1,
              se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(colour = mag >= quantile(mag, 0.5),
                 shape  = mag >= quantile(mag, 0.5)),
             size = 3) +

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|cEOF2| > 50%",
                                "FALSE" = "|cEOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|cEOF2| > 50%",
                                "FALSE" = "|cEOF2| < 50%")) +
  scale_y_continuous("ONI") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  panel_background
```

@@

```{r australia-pp-phase, fig.cap = "(ref:australia-pp-phase-cap)", fig.width = width_column, fig.height=3}
gdata <- cmap %>% 
  .[lon %between% box_australia$lon  & lat %between% box_australia$lat] %>% 
  .[, .(pp = weighted.mean(pp, cos(lat*pi/180))), by = time] %>% 
  .[, pp := Anomaly(pp), by = season(time)] %>% 
  .[ceof$eof[[1]]$left, on  = "time"] %>% 
  .[cEOF == "cEOF2"] %>% 
  .[, arg := Arg(hgt)] %>% 
  .[, mag := Mod(hgt)]

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) + I(cos(x)) -1,
              se = FALSE, color = "black", size = 0.5, 
              fullrange = TRUE) +
  geom_point(aes(colour = mag >= quantile(mag, 0.5),
                 shape  = mag >= quantile(mag, 0.5)),
             size = 3) +

@@

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_y_continuous("Precipitation anomaly") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  theme(plot.subtitle = element_text(size = 8)) +
  panel_background
```

>>>>

theme_set(theme_shceof(base_size = 10) +
            theme(legend.box.spacing = grid::unit(-.5, "lines")))

@@

  geom_qmap(size = 0.1) +
  scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(10))  +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL, 
                   limits = c(NA, -20)) +
  scale_x_longitude() +
  geom_coords() +
  coord_polar() +
  theme(axis.text = element_blank(), 
        legend.box.spacing = grid::unit(-1, "lines")) +
  facet_wrap(.~EOF, ncol = 2)  +
  tagger::tag_facets("rc")

@@

```{r enso-phase, fig.cap = "ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) -1,
              se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(color = q,
                 shape  = q),
             size = 3) +

  scale_color_brewer("", palette = "Dark2") +
  scale_shape_manual("", values = c(20, 18)) +
  scale_y_continuous("ONI") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  panel_background
```

@@

```{r australia-pp-phase, fig.cap = "(ref:australia-pp-phase-cap)", fig.width = width_column, fig.height=3}
gdata <- cmap %>% 
  .[lon %between% box_australia$lon  & lat %between% box_australia$lat] %>% 
  .[, .(pp = weighted.mean(pp, cos(lat*pi/180))), by = time] %>% 
  .[, pp := Anomaly(pp), by = season(time)] %>% 
  .[ceof$eof[[1]]$left, on  = "time"] %>% 
  .[cEOF == "cEOF2"] %>% 
  .[, arg := Arg(hgt)] %>% 
  .[, mag := Mod(hgt)] %>% 
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) + I(cos(x)) -1,
              se = FALSE, color = "black", size = 0.5, 
              fullrange = TRUE) +
  geom_point(aes(colour = q,
                 shape  = q),
             size = 3) +

@@

  scale_color_brewer(NULL, palette = "Dark2") +
  scale_shape_manual(NULL, values = c(20, 18)) +
  scale_y_continuous("Precipitation anomaly") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  theme(plot.subtitle = element_text(size = 8)) +
  panel_background
```
